---
fig-cap-location: margin
reference-location: margin
---

# Cell Modification

```{r}
#| include: false

library(gt)
library(tibble)
```


## siRNA Transfection

[This protocol is adapted from the ambion Ambion Life Technologies protocol (Protocol Pub. No. MAN0007836 Rev. 1.0)]{.aside}

:::{.callout-note}
This protocol is for 6 well plates. For reagent volumes for smaller well form factors, see [here](https://web.archive.org/web/20210709195152/http://tools.thermofisher.com/content/sfs/manuals/Silencer_Select_PreDsignd_Vdtd_CustomDesign_siRNA_man.pdf)
:::

### siRNA Resuspension

1. Briefly centrifuge the tube to ensure dried siRNA is at the bottom of the tube
2. Dilute siRNA down to $100 \mu M$ $(100 \frac{pmoles}{\mu L})$ using nuclease-free water
3. Aliquot out some of the siRNA

:::{.callout-tip}
Each well of a 6-well plate takes $3 \mu L$ of siRNA. Therefore, aliquoting $12 \mu L$ per tube is good enough for triplicates, plus a little extra for the 'angel's share'
:::

4. Store at $\le 20 ^\circ C$


### Performing Transfection

1. Split your cells and dilute them with medium to $200,000\ cells/mL$. You will need $2mL/well$.
2. In 1.5mL eppendorf, prepare the following **per well**:

```{r}
#| echo: false
tibble(reagent = c("Opti-MEM Medium", "Lipofectamine RNAiMAX Reagent"), Volume = c("150uL", "9uL"), `Ref #` = c("31985062", "13778150")) |> 
  gt(rowname_col = "reagent")
```

3. In another 1.5mL eppendorf, prepare the following **per well**:

```{r}
#| echo: false
tibble(reagent = c("Opti-MEM Medium", "siRNA (10 uM)"), Volume = c("150uL", "3uL (30pmol)"), `Ref #` = c("31985062", "varies")) |> 
  gt(rowname_col = "reagent")
```

4. Pipette to mix each tube

5. Mix siRNA and lipofectamine together in a 1:1 ratio (for a single well, this would be $150 \mu L$ of each together. You can simply add the siRNA solution to your lipofectamine tube)

6. Wait $5\ minutes$. During this time, dispense $2mL$ of your cell suspension in each well of a 6 well plate.

7. Add $250 \mu L$ of mixture to each well.

8. Gently swirl and rock the plate to ensure even mixing of cells and siRNA particles

9. Incubate for $3\ days$. There is no need to change medium.

## Inducible shRNA

[This protocol is based of the Horizon Dharmacon SMARTvector Inducible Lentiviral shRNA technical manual]{.aside}

Before you begin, ensure that your cell line has not already been optimized.

### Optimized cell line data

```{r}
#| echo: false
tibble(`Cell Line` = c("5637", "UC9", "UC6", "RT112", "UC14"),
       `Seeding density (cells/100uL)` = c(7500, 5000, 1000, 3000, NA),
       `Need serum?` = c(NA, NA, F, F, NA),
       `Max polybrene (ug/mL)` = c(NA, NA, ">14", ">14", NA), 
       `Relative transduction efficiency` = c(NA, NA, 0.43, NA, NA),
       `Optimal puromycin (ug/mL)` = c(NA, NA, 2.5, 2.5, 10),
       `Date` = c("2021-03-17", "2021-03-17", "2021-04-28", "2021-04-28", "2022-05-19"),
       `Operator` = "Kai Aragaki") |> 
  gt()

```


### Optimizing promoter

Generally, we presume the optimal promoter to be PGK. If you find you are getting poor expression of GFP and poor knockdown despite most of the population being positive, you may try performing promoter optimization for this particular cell line.


### Optimizing transfection conditions





### Determining transduction efficiency




## CRISPR
