---
execute:
  echo: false
---

# FFPE Extraction Methods

```{r}
library(gt)
library(tibble)
```


:::{.callout-note}
Mix solutions well before using

Flick and spin down any tiny reagents before using.
:::


# Sample Preparation

## Slides (Macrodissection)

:::{.callout-note}
You'll need:

  - FFPE slides and _marked_ H+E slides  
  - Scalpel blades 
:::

1. Line up your unstained slide to the marked H+E slide, with the H+E slide on the bottom
2. Trace the area of interest with your scalpel blade
3. Carefully scrape off just the area of interest
4. Transfer the scrapings into the cap of a 1.5mL microfuge tube and press them together
5. Transfer the compressed scrapings into the body of the tube
6. Store scrapings at $-20^\circ C$


### Cores

1. Crush cores with a blue pestle in a 1.5mL microfuge tube. 

:::{.callout-tip}
It helps to position to core on the lower inclined section of the tube, a bit above the bottom.
:::

After RNA (and DNA) isolation, QC the samples using the Nanodrop, Tapestation, and Qubit.

# RNA

## High Pure miRNA Isolation Kit - Roche Kit (Cat. #05080576001)

:::{.callout-note}
* Get ice
* Set heatblock to $55^\circ C$ 
* Set heatblock to $37^\circ C$
* Prepare proteinase K by reconstituting the powder in  
$4.5 mL$ Elution Buffer (Store at $-20^\circ C$) OR thaw previously prepared proteinase K on ice before beginning 
:::

### Deparaffinization with Xylenes

1. Add $800 \mu L$ of xylenes to the microfuge tube with the prepared tissue. Mix by inverting.
2. Incubate for $5\ minutes$ at $room\ temperature$
3. Add $400 \mu L$ of ethanol, vortex.
4. Centrifuge for $2\ minutes$ at $20,000 \times g$
5. Carefully remove supernatant without touching the pellet

:::{.callout-warning}
The supernatant for this step and the following step needs to be disposed of in a bottle specifically labeled to be xylenes and ethanol waste. This is found in the hood space in the first bay.
:::

:::{.callout-warning}
For samples that are super tiny (i.e. can't see a pellet after spinning down), leave about $10-20\mu L$ of the supernatant in at this step. You don't want to risk removing the sample with the supernatant.
:::

6. Add $1 mL$ of ethanol, vortex
7. Centrifuge for $2\ minutes$ at $20,000 \times g$
8. Carefully remove the supernatant without touching the pellet, take note of properly discarding supernatant
9. Dry pellet at $55^\circ C$ for $15\ minutes$ or until all the residual ethanol has dried.
10. Add $100 \mu L$ Paraffin Tissue Lysis Buffer, $16 \mu L$ 10% SDS, and $40 \mu L$ Proteinase K
11. Vortex until pellet has broken up
12. Incubate for $3\ hours$ at $55^\circ C$
13. Place on ice for $3\ minutes$
14. Centrifuge for $15\ minutes$ at $20,000 \times g$
15. Transfer the RNA-containing supernatant to a clean tube
16. Store the pellet at $-20^\circ C$ for later DNA isolation

### RNA Isolation

:::{.callout-note}
For brevity, $spin$ will be shorthand for centrifuging for $30\ seconds$ at $13,000 \times g$
:::

:::{.callout-warning}
Note that this shorthand $spin$ is different than that for the Qiagen kit
:::

:::{.callout-note}
Begin thawing:

- 10X DNase Incubation Buffer  
- DNase I
:::

1. Add $312 \mu L$ Binding Buffer to RNA-containing supernatant, vortex
2. Add $325 \mu L$ Binding Enhancer, vortex
3. Transfer about half of the solution to a High Pure Filter Tube 
4. $Spin$. Discard flow through.
5. Add remaining sample solution.
6. $Spin$. Discard the flow through.
7. Add $500 \mu L$ Wash Buffer to spin column
8. $Spin$. Discard the flow through.
9. Repeat step 8.
10. Add $300 \mu L$ Wash Buffer to spin column
11. $Spin$. Discard the flow through
12. Centrifuge for $1\ minute$ at $13,000 \times g$
13. Place filter in *new* 1.5 mL microfuge tube
14. Add $100 \mu L$ Elution Buffer to 
filter
15. Incubate for $1\ minute$ at $RT$

:::{.callout-warning}
Avoid touching the pipette to the filter or the sides of the tube.
:::

16. Centrifuge for $1\ minute$ at $13,000 \times g$

:::{.callout-tip}
This can be a stopping point - store intermediate RNA at $-80^\circ C$ until you can continue with protocol. 
:::

:::{.callout-note}
Create DNase Solution.  
For $n$ samples:

  - $49.5 \mu L \times\ n$ Nuclease Free Water  
  - $19.5 \mu L \times\ n$ 10X DNase Incubation Buffer  
  - $1 \mu L \times\ n$ DNase I
:::

17. Add $56 \mu L$ DNase Solution, vortex
18. Incubate for $30\ minutes$ at $37^\circ C$
19. Add $325 \mu L$ Binding Buffer 
20. Add $210 \mu L$ Binding Enhancer
21. Transfer whole solution to a *new* High Pure Filter Tube
22. $Spin$. Discard the flow through
23. Repeat step 22.
24. Add $500 \mu L$ Wash Buffer to spin column
25. $Spin$. Discard the flow through.
26. Repeat step 25.
27. Add $300 \mu L$ Wash Buffer to spin column
28. $Spin$. Discard the flow through 
29. Centrifuge for $1\ minute$ at $13,000 \times g$
30. Place filter in *new* 1.5mL microfuge tube, add $40 \mu L$ Nuclease Free Water/Elution Buffer directly to the filter

:::{.callout-warning}
Avoid touching the pipette to the filter or the sides of the tube.
:::

31. Incubate for $1\ minute$ at $RT$
32. Centrifuge for $1\ minute$ at $13,000 \times g$, discard filter
33. The resultant fluid is your isolated RNA. Store at $-80^\circ C$

## Allprep FFPE Kit - Qiagen Kit (Cat. #80234)

:::{.callout-note}
* Get ice
* Set heatblock to $37^\circ C$ 
* Set heatblock to $56^\circ C$
* Set heatblock to $80^\circ C$
* Get RNeasy MinElute (pink) spin column from $4^\circ C$ fridge.
:::


### Deparaffinization of FFPE Tissue with Xylene

1. Add $1 mL$ of xylene to the tube containing your tissue, vortex
2. Centrifuge for $2\ minutes$ at $20,000 \times g$.

:::{.callout-note}
You want the tissue to break up during this and the following steps.
:::

3. Carefully remove the supernatant without disturbing the pellet.

:::{.callout-warning}
The supernatant for this step and the following step needs to be disposed of in a bottle specifically labeled for xylenes and ethanol waste. This 
is usually found in the hood space in the first bay.
:::

4. Add $1 mL$ of ethanol, vortex vigorously, centrifuge for $2\ minutes$ at $20,000 \times g$.
5. Carefully remove the supernatant without disturbing the pellet.
6. Invert the tube briefly to get rid of excess ethanol.

:::{.callout-tip}
I usually have a kimwipe under my tube rack, and lean the tubes upside down against the rack.
:::

7. Dry pellet for $15\ minutes$ at $37^\circ C$ or until the tissue has dried.
8. Re-suspend the pellet in $150 \mu L$ Buffer PKD and $10 \mu L$ Proteinase K, vortex.
9. Incubate for $15\ minutes$ at $56^\circ C$.
10. Incubate on ice for $3\ minutes$.
11. Centrifuge for $15\ minutes$ at $20,000 \times g$.
12. Carefully transfer the RNA-containing supernatant to a new tube. The pellet will be used for DNA isolation later and should be stored at $-20^\circ C$.

### RNA Isolation

:::{.callout-note}
For brevity, $spin$ will be shorthand for centrifuging for $1\ minute$ at $14,000 \times g$
:::

:::{callout-warning}
Note that this shorthand $spin$ is different than that for the Roche kit
:::

:::{.callout-note}
Take Buffer RDD out of the $4^\circ C$ fridge
:::

1. Incubate supernatant for $15\ minutes$ at $80^\circ C$.
2. Add $320 \mu L$ Buffer RLT, vortex.
3. Add $720 \mu L$ ethanol, vortex.
4. Transfer $~500 \mu L$ of mixture to RNease MinElute spin column (pink).
5. $Spin$, discard flow through, $spin$ again.
6. Repeat step 5 with any additional material.
7. $Spin$ to dry filter
8. Put filter in a new collection tube.
9. Add $350 \mu L$ Buffer FRN to spin column.
10. $Spin$, discard flow through. 
11. $Spin$ again.
12. $Spin$ to dry filter. 
13. Put filter in a new collection tube.

:::{.callout-note}
Create DNase I Incubation mix.  
For $n$ samples:

- $10 uL \times n$ DNase I Stock  
- $70 uL \times n$ Buffer RDD  
:::

14. Add $80 \mu L$ of DNase I Incubation Mix to spin column membrane.
15. Incubate at $RT$ for $15\ minutes$.
16. Add $500 \mu L$ Buffer FRN to spin column.
17. $Spin$. *Save the flow through from this step.*
18. Add spin column to a new collection tube and transfer flow through from previous step to spin column.
19. $Spin$, discard flow through, 
20. $Spin$ again
21. $Spin$ to dry filter, put filter in a new collection tube.
22. Add $500 \mu L$ Buffer RPE to spin column.
23. $Spin$, discard flow through *(x3)*.
24. Add $500 \mu L$ Buffer RPE to spin column.
25. $Spin$, discard flow through *(x3)*.
26. Place filter column into fresh collection tube.
27. $Spin$ for $5\ minutes$ with lids open.
28. Place filter column into fresh 1.5 mL microfuge tube
29. Add $30 \mu L$ Nuclease Free Water to center of filter column.
30. Incubate for $1\ minute$ at $RT$
31. $Spin$.

:::{.callout-tip}
Avoid touching the walls of the filter column when pipetting in the water. 
:::

32. Store RNA at $-80^\circ C$.  Use Qiagen Allprep FFPE (cat. 80234) for DNA isolation.

## MagMax FFPE RNA/DNA Ultra Kit (Cat. #A31881)

With this protocol, you isolated DNA and RNA on the same day, unlike the previous protocols.

:::{.callout-note}
- Get ice
- Set heatblock to $55^\circ C$
- Set heatblock to $90^\circ C$
- Take out Protease
- Locate the rack magnet
:::

:::{.callout-tip}
This protocol involves using different volumes/concentrations of reagents depending on starting tissue size. It may help to do similar sized tissue at once.
:::

Before beginning, calculate the volumes of reagents you will need based on the following table.  It will help to do $(n+1)x$, where $n$ is the number of samples and $x$ is the volume of the reagent.

Protease Solution:
```{r echo = F}
tibble("Reagent" = c("Protease", "Proteaste Digestion Buffer", "Total Volume"),
       "Section <= 40 um" = c("10 uL", "100 uL", "110 uL"),
       ">40 um" = c("10 uL", "200 uL", "210 ul")) |> 
  gt(rowname_col = "Reagent")
```

DNA Binding Buffer:
```{r echo = F}
tibble("Reagent" = c("Binding Solution", "Nucleic Acid Binding Beads", "Total Volume"),
       "Section <= 40 um" = c("135 uL", "20 uL", "155 uL"),
       "> 40 um" = c("250 uL", "20 uL", "270 uL")) |> 
  gt(rowname_col = "Reagent")
```

The following two solutions are not dependent on starting tissue size.
RNA Binding Buffer:
```{r echo = F}
tibble("Reagent" = c("Binding Solution", "Isopropanol", "Total Volume"),
       "Section <= 40 um" = c("100 uL", "300 uL", "400 uL"),
       "> 40 um" = c("150 uL", "500 uL", "650 uL")) |> 
  gt(rowname_col = "Reagent")
```

DNase Solution:
```{r echo = F}
tibble("Reagent" = c("DNase", "DNase Buffer", "Nuclease-Free Water", "Total Volume"),
       "Volume" = c("20 uL", "10 uL", "70 uL", "100 uL")) |> 
  gt(rowname_col = "Reagent")
```

RNA Rebinding Buffer:
```{r echo = F}
tibble("Reagent" = c("Binding Solution", "Isopronanol", "Total Volume"),
       "Volume" = c("200 uL", "250 uL", "450 uL")) |> 
  gt(rowname_col = "Reagent")
```

#### Deparaffinization of FFPE Tissue with Xylene

The deparaffinization of the original protocol has been replaced with the deparaffinization of the Roche protocol. It just works a little bit better and is easier for the technician.

1. Add $800 \mu L$ of xylenes, mix by flipping tube upside down, 
incubate for $5\ minutes$ at $RT$
2. Add $400 \mu L$ of ethanol, vortex, centrifuge for $2\ minutes$ at $20,000 \times g$
3. Carefully remove supernatant without touching the pellet

:::{.callpout-warning}
The supernatant for this step and the following step needs to be disposed of in a bottle specifically labeled to be xylenes and ethanol waste. This is usually found in the hood space in the first bay.
:::

:::{.callout-tip}
For samples that are super tiny (i.e. can't see a pellet after spinning down), leave about 10-20 uL of the supernatant in at this step. You don't want to risk removing the sample with the supernatant.
:::

4. Add $1 mL$ of ethanol, vortex
5. Centrifuge for $2\ minutes$ at $20,000 \times g$
6. Carefully remove the supernatant without touching the pellet, take note of properly discarding supernatant
7. Dry pellet for $15\ minutes$ at $55^\circ C$ or until all the residual ethanol has dried.
8. Add Protease Solution to the dried tissue according to the table:

```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Volume" = c("110 uL", "210 uL")) |> 
  gt()
```
9. Gently vortex to make sure the tissue becomes dislodged from the tube. 
10. Incubate at $55^\circ C$ for $1\ hour$.
11. Incubate at $90^\circ C$ for $1\ hour$. Let cool before moving on.

:::{.callout-note}
Take out the Nucleic Acid Binding Beads.
:::

#### DNA Isolation

:::{.callout-note}
Briefly spin down samples after mixing.
:::

1. Add DNA Binding Buffer (pipette slowly) according to the table:

```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Volume" = c("155 uL", "270 uL")) |> 
  gt()
```

Precipitant may form at this step, but does not interfere with DNA binding.

2. Mix for $5\ minutes$ according to the table:
```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Speed" = c("1000 rpm", "1150 rpm")) |> 
  gt()
```

:::{.callout-tip}
For the vortexing steps, use the 96 well plate vortex. You can move it to your bench. I've marked 2 little black lines, the first one at ~1000 rpm, the second at ~1150 rpm. You'll need to secure the tubes while you vortex.
:::

3. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears. There should be a "smear" of beads on one side of the tube.
4. Carefully transfer the RNA containing supernatant to a fresh labelled tube.
5. Wash the beads with DNA Wash Buffer according to the table:

```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Volume" = c("200 uL", "400 uL")) |> 
  gt()
```


6. Mix for $1-2\ minutes$ at $1100\ rpm$.
7. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
8. Carefully remove the supernatant without disturbing the pellet.
9. Repeat steps 16-19.
10. Wash the beads with Wash Solution 2 (will need ethanol added, please see bottle for instructions) according to the table:

```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Volume" = c("200 uL", "500 uL")) |> 
  gt()
```

11. Mix for $1\ minute$ at $1150\ rpm$.
12. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
13. Carefully remove the supernatant without disturbing the pellet.
14. Repeat steps 21-24.
15. Let beads airdry (top of caps open) for $3\ minutes$. It may help to go in with a smaller pipette and remove any drops that may be left over.
16. Add $50 \mu L$ Elution Buffer to the beads.
17. Mix for $5\ minutes$ at $1150\ rpm$.
18. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
19. Carefully move the supernatant to a clean labelled tube.  This is your DNA.
20. Store at $-20^\circ C$.

#### RNA Isolation

:::{.callout-note}
- Take out DNase and DNase buffer.
- Set heatblock to $37^\circ C$.
:::

1. Add $20 \mu L$ Nucleic Acid Binding Beads to the RNA containing supernatant.
2. Add RNA Binding Buffer according to the table:
```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Volume" = c("400 uL", "650 uL")) |> 
  gt()
```

3. Mix for $5\ minutes$ at $1150\ rpm$.
4. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
5. Carefully discard the supernatant without disturbing the pellet.
6. Wash the beads with $500 \mu l$ RNA Wash Buffer.
7. Mix for $1\ minute$ at $1150\ rpm$.
8. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
9. Carefully discard the supernatant without disturbing the pellet.
10. Wash the beads with $500 \mu L$ Wash Solution 2.
11. Mix for $1\ minute$ at $1150\ rpm$.
12. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
13. Carefully discard the supernatant without disturbing the pellet.
15. Let beads airdry (top of caps open) for $3\ minutes$.  It may help to go in with a smaller pipette and remove any drops that may be left over.
16. Add $100 \mu L$ DNase Solution.
17. Incubate at $37^\circ C$ for $20\ minutes$.
18. Add $450 \mu L$ RNA Rebinding Buffer.
19. Mix for $5\ minutes$ at $1150\ rpm$.
20. Place the samples in the magnetic rack for $5\ minutes$ until the solution clears.
21. Carefully discard the supernatant without disturbing the pellet.
22. Wash beads with $500 \mu L$ RNA Wash Buffer.
23. Mix for $1\ minute$ at $1150\ rpm$.
24. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
25. Carefully discard the supernatant without discarding the pellet.
26. Wash the beads with $500 \mu L$ Wash Solution 2.
27. Mix for $1\ minute$ at $1150\ rpm$.
28. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
29. Carefully discard the supernatant without discarding the pellet.
30. Repeat steps 26-29.
31. Let beads airdry (top of caps open) for $3\ minutes$. It may help to go in with a smaller pipette and remove any drops that may be left over.
32. Add $50 \mu L$ Elution Solution.
33. Mix for $5\ minutes$ at $1150\ rpm$.
34. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
35. Carefully move the supernatant to a clean labelled tube. This is your RNA.
36. Store RNA at $-80^\circ C$.

# DNA

There are two different Qiagen FFPE DNA isolation kits that we use - Qiagen Allprep FFPE (cat. 80234) and Qiagen DNA FFPE (cat. 56404).  

## Qiagen Allprep FFPE (Cat. #80234)

:::{.callout-note}
For brevity, $spin$ will be shorthand for centrifuging for $1\ minute$ at $16,000 \times g$
:::

:::{.callout-warning}
Note that this shorthand is different than those above!
:::

:::{.callout-note}
* Set heatblock to $56^\circ C$.
* Set heatblock to $90^\circ C$.
* Get QIAamp MinElute column (white) from $4^\circ C$.
:::

1. Re-suspend pellet in $180 \mu L$ Buffer ATL and $20 \mu L$ Proteinase K. Vortex.
2. Incubate at $56^\circ C$ for $1\ hour$.
3. Incubate at $90^\circ C$ for $2\ hours$.

:::{.callout-warning}
Place an unused heating block on the tube lids to keep them from opening during incubation.
:::

4. Briefly centrifuge to remove any drops from the lid.
5. Add $20 \mu L$ Rnase A to the sample, mix.
6. Incubate for $2 minutes$ at $RT$.
7. Add $200 \mu L$ Buffer AL to the sample, vortex.  You will see precipitants form.
8. Add $200 \mu L$ ethanol to the sample, vortex. Spin down.
9. Transfer the sample to a QIAamp MinElute Column (white).
10. $Spin$. Discard flow through.
11. $Spin$.
12. Repeat steps 9-11 with any leftover sample.
13. Add $650 \mu L$ Buffer AW1 to spin column.
14. $Spin$. Discard flow through. (x2)
15. Add $650 \mu L$ Buffer AW2 to spin column.
16. Repeat step 14.
17. Add $650 \mu L$ ethanol to spin column.
18. Repeat step 14.
19. Place spin column in a new collection tube.
20. $Spin$ for $5\ minutes$ with lids open.
21. Place column into labelled 1.5 mL tube.
22. Add $60 \mu L$ Nuclease Free Water.
23. Incubate for $1\ minute$ at $RT$.
24. $Spin$.
25. The flow-through is your DNA - store at $-20^\circ C$.


## Qiagen DNA FFPE (Cat. #56404)

:::{.callout-note}
For brevity, $spin$ will be shorthand for centrifuging for $1\ minute$ at $16,000 \times g$
:::

:::{.callout-note}
* Set heatblock to $56^\circ C$.
* Set heatblock to $90^\circ C$.
* Get QIAamp MinElute column (white) from $4^\circ C$.
:::

1. Re-suspend pellet in $180 \mu L$ Buffer ATL and $20 \mu L$ Proteinase K. Vortex.
2. Incubate at $56^\circ C$ for $1\ hour$.
3. Incubate at $90^\circ C$ for $2\ hours$.

:::{.callout-warning}
Place an unused heating block on the tubes lids to keep them from opening during incubation.
:::

4. Briefly centrifuge to remove any drops from the lid.
5. Add $20 \mu L$ Rnase A to the sample, vortex.
6. Incubate for $2\ minutes$ at $RT$.
7. Add $200 \mu L$ Buffer AL to the sample, vortex. You will see precipitants form.
8. Add $200 \mu L$ ethanol to the sample, vortex. Spin down.
9. Transfer the sample to a QIAamp MinElute Column (white).
10. $Spin$. Discard flow through.
11. $Spin$.
12. Repeat steps 9-11 with any leftover sample.
13. Add $500 \mu L$ Buffer AW1 to spin column.
14. $Spin$. Discard flow through. (x2)
15. Add $500 \mu L$ Buffer AW2 to spin column.
16. Repeat step 14.
19. Place spin column in a new collection tube.
20. $Spin$ for $3\ minutes$ with lids open.
21. Place column into labelled 1.5 mL tube.
22. Add $60 \mu L$ Nuclease Free Water.
23. Incubate for $1\ minute$ at $RT$.
24. $Spin$.
25. The flow-through is your DNA - store at $-20^\circ C$.
