# FFPE Extraction Methods

```{r}
library(gt)
library(tibble)
```


:::info
Mix solutions well before using
:::

:::info
Flick and spin down any tiny reagents before using.
:::


## Sample Preparation

### Slides (Macrodissection)

:::prep
Ensure you have:

  - FFPE slides and _marked_ H+E slide  
  - Scalpel blade  
:::

1. Line up your unstained slide to the marked H+E slide, with the H+E slide on the bottom
2. Trace the area of interest with your scalpel blade
3. Carefully scrape off just the area of interest

:::tip
It's easier to carefully put the tissue in a 1.5mL tube _cap_, and try to smush it all together, _then_ move it to the tube body.
:::

:::tip
Store scrapings at $-20^\circ C$.
:::

### Cores

:::prep
Ensure you have:

  - FFPE cores  
  - Blue mortar (to crush the cores)  
:::

To prepare the cores, crush them with a blue mortar in a 1.5mL microfuge tube. 

:::tip
It helps to position to core on the lower inclined section of the tube, a bit above the bottom.
:::

After RNA (and DNA) isolation, QC the samples using the Nanodrop, Tapestation, and Qubit.

## RNA

### High Pure miRNA Isolation Kit - Roche Kit (Cat. #05080576001)

:::prep
* Get ice
* Set heatblock to $55^\circ C$ 
* Set heatblock to $37^\circ C$
* Prepare proteinase K by reconstituting the powder in  
$4.5 mL$ Elution Buffer (Store at $-20^\circ C$) OR thaw pre-prepared proteinase K on ice before beginning 
:::

#### Deparaffinization using Xylenes

1. Add $800 \mu L$ of __xylenes__ to the microfuge tube with the prepared tissue. Mix by inverting.
2. Incubate for $5\ minutes$ at $RT$
3. Add $400 \mu L$ of __ethanol__, vortex.
4. Centrifuge for $2\ minutes$ at $20,000 \times g$
5. Carefully remove supernatant without touching the pellet

:::warn
The supernatant for this step and the following step needs to be disposed of in a bottle specifically labeled to be xylenes and ethanol waste. This is found in the hood space in the first bay.
:::

:::tip
For samples that are super tiny (i.e. can't see a pellet after spinning down), leave about 10-20 <span>$\mu L$<span> of the supernatant in at this step. You don't want to risk removing the sample with the supernatant.
:::

6. Add $1 mL$ of __ethanol__, vortex
7. Centrifuge for $2\ minutes$ at $20,000 \times g$
8. Carefully remove the supernatant without touching the pellet, take note of properly discarding supernatant
9. Dry pellet at $55^\circ C$ for $15\ minutes$ or until all the residual ethanol has dried.
10. Add $100 \mu L$ __Paraffin Tissue Lysis Buffer__, $16 \mu L$ __10% SDS__, and $40 \mu L$ __Proteinase K__
11. Vortex until pellet has broken up
12. Incubate for $3\ hours$ at $55^\circ C$
13. Place on ice for $3\ minutes$
14. Centrifuge for $15\ minutes$ at $20,000 \times g$
15. Transfer the RNA-containing supernatant to a clean tube
16. Store the pellet at $-20^\circ C$ for later DNA isolation

#### RNA Isolation

:::info
For brevity, $spin$ will be shorthand for centrifuging for $30\ seconds$ at $13,000 \times g$
:::

:::warn
Note that this shorthand $spin$ is different than that for the Qiagen kit
:::

:::prep
Begin thawing:

- 10X DNase Incubation Buffer  
- DNase I
:::

1. Add $312 \mu L$ __Binding Buffer__ to RNA-containing supernatant, vortex
2. Add $325 \mu L$ __Binding Enhancer__, vortex
3. Transfer about half of the solution to a __High Pure Filter Tube__ 
4. $Spin$. Discard flow through.
5. Add remaining sample solution.
6. $Spin$. Discard the flow through.
7. Add $500 \mu L$ __Wash Buffer__ to spin column
8. $Spin$. Discard the flow through.
9. Repeat step 8.
10. Add $300 \mu L$ __Wash Buffer__ to spin column
11. $Spin$. Discard the flow through
12. Centrifuge for $1\ minute$ at $13,000 \times g$
13. Place filter in *new* 1.5 mL microfuge tube
14. Add $100 \mu L$ __Elution Buffer__ to 
filter
15. Incubate for $1\ minute$ at $RT$

:::warn
Be careful to not touch pipette tip to the filter or the sides of the tube.
:::

16. Centrifuge for $1\ minute$ at $13,000 \times g$

:::tip
This can be a stopping point - store intermediate RNA at $-80^\circ C$ until you can continue with protocol. 
:::

:::prep
Create DNase Solution.  
For $n$ samples:

  - $49.5 \mu L \times\ n$ Nuclease Free Water  
  - $19.5 \mu L \times\ n$ 10X DNase Incubation Buffer  
  - $1 \mu L \times\ n$ DNase I
:::

17. Add $56 \mu L$ __DNase Solution__, vortex
18. Incubate for $30\ minutes$ at $37^\circ C$
19. Add $325 \mu L$ __Binding Buffer__ 
20. Add $210 \mu L$ __Binding Enhancer__
21. Transfer whole solution to a *new* High Pure Filter Tube
22. $Spin$. Discard the flow through
23. Repeat step 22.
24. Add $500 \mu L$ __Wash Buffer__ to spin column
25. $Spin$. Discard the flow through.
26. Repeat step 25.
27. Add $300 \mu L$ __Wash Buffer__ to spin column
28. $Spin$. Discard the flow through 
29. Centrifuge for $1\ minute$ at $13,000 \times g$
30. Place filter in *new* 1.5mL microfuge tube, add $40 \mu L$ __Nuclease Free Water/Elution Buffer__ directly to the filter

:::warn
Avoid touching the pipette to the filter or the sides of the tube.
:::

31. Incubate for $1\ minute$ at $RT$
32. Centrifuge for $1\ minute$ at $13,000 \times g$, discard filter
33. The resultant fluid is your isolated RNA. Store at $-80^\circ C$

### Allprep FFPE Kit - Qiagen Kit (Cat. #80234)

:::prep
* Get ice
* Set heatblock to $37^\circ C$ 
* Set heatblock to $56^\circ C$
* Set heatblock to $80^\circ C$
* Get RNeasy MinElute (pink) spin column from $4^\circ C$ fridge.
:::


#### Deparaffinization of FFPE Tissue with Xylene

1. Add $1 mL$ of __xylene__ to the tube containing your tissue, vortex
2. Centrifuge for $2\ minutes$ at $20,000 \times g$.

:::info
You want the tissue to break up during this and the following steps.
:::

3. Carefully remove the supernatant without disturbing the pellet.

:::warn
The supernatant for this step and the following step needs to be disposed of in a bottle specifically labeled for xylenes and ethanol waste. This 
is usually found in the hood space in the first bay.
:::

4. Add $1 mL$ of __ethanol__, vortex vigorously, centrifuge for $2\ minutes$ at $20,000 \times g$.
5. Carefully remove the supernatant without disturbing the pellet.
6. Invert the tube briefly to get rid of excess ethanol.

:::tip
I usually have a kimwipe under my tube rack, and lean the tubes upside down against the rack.
:::

7. Dry pellet for $15\ minutes$ at $37^\circ C$ or until the tissue has dried.
8. Re-suspend the pellet in $150 \mu L$ __Buffer PKD__ and $10 \mu L$ __Proteinase K__, vortex.
9. Incubate for $15\ minutes$ at $56^\circ C$.
10. Incubate on ice for $3\ minutes$.
11. Centrifuge for $15\ minutes$ at $20,000 \times g$.
12. Carefully transfer the RNA-containing supernatant to a new tube. The pellet will be used for DNA isolation later and should be stored at $-20^\circ C$.

#### RNA Isolation

:::info
For brevity, $spin$ will be shorthand for centrifuging for $1\ minute$ at $14,000 \times g$
:::

:::warn
Note that this shorthand $spin$ is different than that for the Roche kit
:::

:::prep
Take Buffer RDD out of the $4^\circ C$ fridge
:::

1. Incubate supernatant for $15\ minutes$ at $80^\circ C$.
2. Add $320 \mu L$ __Buffer RLT__, vortex.
3. Add $720 \mu L$ __ethanol__, vortex.
4. Transfer $~500 \mu L$ of mixture to __RNease MinElute spin column__ (pink).
5. $Spin$, discard flow through, $spin$ again.
6. Repeat step 5 with any additional material.
7. $Spin$ to dry filter
8. Put filter in a new collection tube.
9. Add $350 \mu L$ __Buffer FRN__ to spin column.
10. $Spin$, discard flow through. 
11. $Spin$ again.
12. $Spin$ to dry filter. 
13. Put filter in a new collection tube.

:::prep
Create DNase I Incubation mix.  
For $n$ samples:

- $10 uL \times n$ DNase I Stock  
- $70 uL \times n$ Buffer RDD  
:::

14. Add $80 \mu L$ of __DNase I Incubation Mix__ to spin column membrane.
15. Incubate at $RT$ for $15\ minutes$.
16. Add $500 \mu L$ __Buffer FRN__ to spin column.
17. $Spin$. *Save the flow through from this step.*
18. Add spin column to a new collection tube and transfer flow through from previous step to spin column.
19. $Spin$, discard flow through, 
20. $Spin$ again
21. $Spin$ to dry filter, put filter in a new collection tube.
22. Add $500 \mu L$ __Buffer RPE__ to spin column.
23. $Spin$, discard flow through *(x3)*.
24. Add $500 \mu L$ __Buffer RPE__ to spin column.
25. $Spin$, discard flow through *(x3)*.
26. Place filter column into fresh collection tube.
27. $Spin$ for $5\ minutes$ with lids open.
28. Place filter column into fresh 1.5 mL microfuge tube
29. Add $30 \mu L$ __Nuclease Free Water__ to center of filter column.
30. Incubate for $1\ minute$ at $RT$
31. $Spin$.

:::tip
Avoid touching the walls of the filter column when pipetting in the NF water. Try to get it directly on the filter.
:::

32. Store RNA at $-80^\circ C$.  Use Qiagen Allprep FFPE (cat. 80234) for DNA isolation.

### MagMax FFPE RNA/DNA Ultra Kit (Cat. #A31881)

With this protocol, you isolated DNA and RNA on the same day, unlike the previous protocols.

:::prep
- Get ice
- Set heatblock to $55^\circ C$
- Set heatblock to $90^\circ C$
- Take out __Protease__
- Locate the rack magnet
:::

:::tip
This protocol involves using different volumes/concentrations of reagents depending on starting tissue size. It may help to do similar sized tissue at once.
:::

Before beginning, calculate the volumes of reagents you will need based on the following table.  It will help to do $(n+1)x$, where $n$ is the number of samples and $x$ is the volume of the reagent.

__Protease Solution__:
```{r echo = F}
tibble("Reagent" = c("Protease", "Proteaste Digestion Buffer", "Total Volume"),
       "Section <= 40 um" = c("10 uL", "100 uL", "110 uL"),
       ">40 um" = c("10 uL", "200 uL", "210 ul")) |> 
  gt(rowname_col = "Reagent")
```

__DNA Binding Buffer__:
```{r echo = F}
tibble("Reagent" = c("Binding Solution", "Nucleic Acid Binding Beads", "Total Volume"),
       "Section <= 40 um" = c("135 uL", "20 uL", "155 uL"),
       "> 40 um" = c("250 uL", "20 uL", "270 uL")) |> 
  gt(rowname_col = "Reagent")
```

The following two solutions are not dependent on starting tissue size.
__RNA Binding Buffer__:
```{r echo = F}
tibble("Reagent" = c("Binding Solution", "Isopropanol", "Total Volume"),
       "Section <= 40 um" = c("100 uL", "300 uL", "400 uL"),
       "> 40 um" = c("150 uL", "500 uL", "650 uL")) |> 
  gt(rowname_col = "Reagent")
```

__DNase Solution__:
```{r echo = F}
tibble("Reagent" = c("DNase", "DNase Buffer", "Nuclease-Free Water", "Total Volume"),
       "Volume" = c("20 uL", "10 uL", "70 uL", "100 uL")) |> 
  gt(rowname_col = "Reagent")
```

__RNA Rebinding Buffer__:
```{r echo = F}
tibble("Reagent" = c("Binding Solution", "Isopronanol", "Total Volume"),
       "Volume" = c("200 uL", "250 uL", "450 uL")) |> 
  gt(rowname_col = "Reagent")
```

#### Deparaffinization of FFPE Tissue with Xylene

The deparaffinization of the original protocol has been replaced with the deparaffinization of the Roche protocol. It just works a little bit better and is easier for the technician.

1. Add $800 \mu L$ of __xylenes__, mix by flipping tube upside down, 
incubate for $5\ minutes$ at $RT$
2. Add $400 \mu L$ of __ethanol__, vortex, centrifuge for $2\ minutes$ at $20,000 \times g$
3. Carefully remove supernatant without touching the pellet

:::warn
The supernatant for this step and the following step needs to be disposed of 
in a bottle specifically labeled to be xylenes and ethanol waste. This 
is usually found in the hood space in the first bay.
:::

:::tip
For samples that are super tiny (i.e. can't see a pellet after spinning down), leave about 10-20 uL of the supernatant in at this step. You don't want to risk removing the sample with the supernatant.
:::

4. Add $1 mL$ of __ethanol__, vortex
5. Centrifuge for $2\ minutes$ at $20,000 \times g$
6. Carefully remove the supernatant without touching the pellet, take note of properly discarding supernatant
7. Dry pellet for $15\ minutes$ at $55^\circ C$ or until all the residual ethanol has dried.
8. Add __Protease Solution__ to the dried tissue according to the table:

```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Volume" = c("110 uL", "210 uL")) |> 
  gt()
```
9. Gently vortex to make sure the tissue becomes dislodged from the tube. 
10. Incubate at $55^\circ C$ for $1\ hour$.
11. Incubate at $90^\circ C$ for $1\ hour$. Let cool before moving on.

:::prep
Take out the Nucleic Acid Binding Beads.
:::

#### Isolate DNA

:::info
Quickly spin down samples after mixing.
:::

1. Add __DNA Binding Buffer__ (pipette slowly) according to the table:

```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Volume" = c("155 uL", "270 uL")) |> 
  gt()
```

Precipitant may form at this step, but does not interfere with DNA binding.

2. Mix for $5\ minutes$ according to the table:
```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Speed" = c("1000 rpm", "1150 rpm")) |> 
  gt()
```

:::tip
For the vortexing steps, use the 96 well plate vortex. You can move it to your bench. I've marked 2 little black lines, the first one at ~1000 rpm, the second at ~1150 rpm. You'll need to secure the tubes while you vortex.
:::

3. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears. There should be a "smear" of beads on one side of the tube.
4. Carefully transfer the RNA containing supernatant to a fresh labelled tube.
5. Wash the beads with __DNA Wash Buffer__ according to the table:

```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Volume" = c("200 uL", "400 uL")) |> 
  gt()
```


6. Mix for $1-2\ minutes$ at $1100\ rpm$.
7. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
8. Carefully remove the supernatant without disturbing the pellet.
9. Repeat steps 16-19.
10. Wash the beads with __Wash Solution 2__ (will need ethanol added, please see bottle for instructions) according to the table:

```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Volume" = c("200 uL", "500 uL")) |> 
  gt()
```

11. Mix for $1\ minute$ at $1150\ rpm$.
12. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
13. Carefully remove the supernatant without disturbing the pellet.
14. Repeat steps 21-24.
15. Let beads airdry (top of caps open) for $3\ minutes$. It may help to go in with a smaller pipette and remove any drops that may be left over.
16. Add $50 \mu L$ __Elution Buffer__ to the beads.
17. Mix for $5\ minutes$ at $1150\ rpm$.
18. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
19. Carefully move the supernatant to a clean labelled tube.  This is your DNA.
20. Store at $-20^\circ C$.

#### Isolate RNA

:::prep
- Take out DNase and DNase buffer.
- Set heatblock to $37^\circ C$.
:::

1. Add $20 \mu L$ __Nucleic Acid Binding Beads__ to the RNA containing supernatant.
2. Add __RNA Binding Buffer__ according to the table:
```{r echo = F}
tibble("Section" = c("<= 40 um", ">40 um"),
       "Volume" = c("400 uL", "650 uL")) |> 
  gt()
```

3. Mix for $5\ minutes$ at $1150\ rpm$.
4. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
5. Carefully discard the supernatant without disturbing the pellet.
6. Wash the beads with $500 \mu l$ __RNA Wash Buffer__.
7. Mix for $1\ minute$ at $1150\ rpm$.
8. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
9. Carefully discard the supernatant without disturbing the pellet.
10. Wash the beads with $500 \mu L$ __Wash Solution 2__.
11. Mix for $1\ minute$ at $1150\ rpm$.
12. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
13. Carefully discard the supernatant without disturbing the pellet.
15. Let beads airdry (top of caps open) for $3\ minutes$.  It may help to go in with a smaller pipette and remove any drops that may be left over.
16. Add $100 \mu L$ __DNase Solution__.
17. Incubate at $37^\circ C$ for $20\ minutes$.
18. Add $450 \mu L$ __RNA Rebinding Buffer__.
19. Mix for $5\ minutes$ at $1150\ rpm$.
20. Place the samples in the magnetic rack for $5\ minutes$ until the solution clears.
21. Carefully discard the supernatant without disturbing the pellet.
22. Wash beads with $500 \mu L$ __RNA Wash Buffer__.
23. Mix for $1\ minute$ at $1150\ rpm$.
24. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
25. Carefully discard the supernatant without discarding the pellet.
26. Wash the beads with $500 \mu L$ __Wash Solution 2__.
27. Mix for $1\ minute$ at $1150\ rpm$.
28. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
29. Carefully discard the supernatant without discarding the pellet.
30. Repeat steps 26-29.
31. Let beads airdry (top of caps open) for $3\ minutes$. It may help to go in with a smaller pipette and remove any drops that may be left over.
32. Add $50 \mu L$ __Elution Solution__.
33. Mix for $5\ minutes$ at $1150\ rpm$.
34. Place the samples in the magnetic rack for $2\ minutes$ until the solution clears.
35. Carefully move the supernatant to a clean labelled tube. This is your RNA.
36. Store RNA at $-80^\circ C$.

## DNA

There are two different Qiagen FFPE DNA isolation kits that we use - Qiagen Allprep FFPE (cat. 80234) and Qiagen DNA FFPE (cat. 56404).  

### Qiagen Allprep FFPE (Cat. #80234)

#### DNA Isolation

:::info
For brevity, $spin$ will be shorthand for centrifuging for $1\ minute$ at $16,000 \times g$
:::

:::warn
Note that this shorthand is different than those above!
:::

:::prep
* Set heatblock to $56^\circ C$.
* Set heatblock to $90^\circ C$.
* Get QIAamp MinElute column (white) from $4^\circ C$.
:::

1. Re-suspend pellet in $180 \mu L$ __Buffer ATL__ and $20 \mu L$ __Proteinase K__. Vortex.
2. Incubate at $56^\circ C$ for $1\ hour$.
3. Incubate at $90^\circ C$ for $2\ hours$.

:::tip
When incubating tubes at high temperatures, it's helpful to place a block on the tubes to keep them from popping open during incubation.
:::

4. Briefly centrifuge to remove any drops from the lid.
5. Add $20 \mu L$ __Rnase A__ to the sample, mix.
6. Incubate for $2 minutes$ at $RT$.
7. Add $200 \mu L$ __Buffer AL__ to the sample, vortex.  You will see precipitants form.
8. Add $200 \mu L$ __ethanol__ to the sample, vortex. Spin down.
9. Transfer the sample to a QIAamp MinElute Column (white).
10. $Spin$. Discard flow through.
11. $Spin$.
12. Repeat steps 9-11 with any leftover sample.
13. Add $650 \mu L$ __Buffer AW1__ to spin column.
14. $Spin$. Discard flow through. (x2)
15. Add $650 \mu L$ __Buffer AW2__ to spin column.
16. Repeat step 14.
17. Add $650 \mu L$ __ethanol__ to spin column.
18. Repeat step 14.
19. Place spin column in a new collection tube.
20. $Spin$ for $5\ minutes$ with lids open.
21. Place column into labelled 1.5 mL tube.
22. Add $60 \mu L$ __Nuclease Free Water__.
23. Incubate for $1\ minute$ at $RT$.
24. $Spin$.
25. The flow-through is your DNA - store at $-20^\circ C$.


### Qiagen DNA FFPE (Cat. #56404)

#### DNA Isolation

:::info
For brevity, $spin$ will be shorthand for centrifuging for $1\ minute$ at $16,000 \times g$
:::

:::prep
* Set heatblock to $56^\circ C$.
* Set heatblock to $90^\circ C$.
* Get QIAamp MinElute column (white) from $4^\circ C$.
:::

1. Re-suspend pellet in $180 \mu L$ __Buffer ATL__ and $20 \mu L$ __Proteinase K__. Vortex.
2. Incubate at $56^\circ C$ for $1\ hour$.
3. Incubate at $90^\circ C$ for $2\ hours$.

:::tip
When incubating tubes at high temperatures, it's helpful to place a block on the tubes to keep them from popping open during incubation.
:::

4. Briefly centrifuge to remove any drops from the lid.
5. Add $20 \mu L$ __Rnase A__ to the sample, vortex.
6. Incubate for $2\ minutes$ at $RT$.
7. Add $200 \mu L$ __Buffer AL__ to the sample, vortex. You will see precipitants form.
8. Add $200 \mu L$ __ethanol__ to the sample, vortex. Spin down.
9. Transfer the sample to a QIAamp MinElute Column (white).
10. $Spin$. Discard flow through.
11. $Spin$.
12. Repeat steps 9-11 with any leftover sample.
13. Add $500 \mu L$ __Buffer AW1__ to spin column.
14. $Spin$. Discard flow through. (x2)
15. Add $500 \mu L$ __Buffer AW2__ to spin column.
16. Repeat step 14.
19. Place spin column in a new collection tube.
20. $Spin$ for $3\ minutes$ with lids open.
21. Place column into labelled 1.5 mL tube.
22. Add $60 \mu L$ __Nuclease Free Water__.
23. Incubate for $1\ minute$ at $RT$.
24. $Spin$.
25. The flow-through is your DNA - store at $-20^\circ C$.
