# MTT Assay {#sec-mtt-assay}
```{r}
#| include: false
library(tibble)
library(gt)
library(gp)
library(ggplot2)
```

## Reagents

### MTT

Resuspension of MTT should be performed in the hood, with sterile PBS.

One aliquot of MTT solution can be prepared as follows:

```{r}
#| echo: false

tibble(Reagent = c("MTT", "PBS"), Amount = c("5mg", "1mL"), `Cat#` = c("114-056-101CS (Core Store)", "298-93-1 (ThermoFisher)")) |> gt()
```

Generally, you should resuspend a larger amount of MTT (final volume 20-30mL) and aliquot it out into labelled 1mL aliquots. These can be frozen and stored at $-20^\circ C$

## Overview


```{r}
#| echo: false
#| label: mtt-layout
#| fig-cap: "Typical MTT layout. Shapes represent different drugs, color represents concetration."
#| column: margin

gp(8, 12) |> 
  gp_sec("Drug", 4, 6, labels = LETTERS[1:4]) |> 
  gp_sec("conc", 4, 1, labels = c(0, 1, 10, 100, 1000, 10000)) |> 
  gp_serve() |> 
  ggplot(aes(.col, .row, color = log10(as.numeric(as.character(conc)) + 1), shape = Drug)) + 
  geom_point(size = 7) + 
  scale_y_reverse() +
  labs(color = "log10([Drug])") + 
  gp::gp_mini_theme()
```

A typical MTT assay is carried out in a 96-well plate^[Cat# 100027, JHU Supply Store]. The concentrations of drug you use depends on the $IC_{50}$ of the drug. This may require a bit of optimization. Your lowest condition should always be DMSO only, with a volume of DMSO equal to your highest-DMSO condition^[This is usually your highest drug condition]. The volume of DMSO should generally not exceed 0.1%^[That is, $\frac{1\ vol}{1000\ vol}$ ], and preferably is less than that (0.01% is great). Generally, we do 10-fold dilutions of the drug (ie 10uM, 1uM...)

:::{.callout-tip}
With this protocol, if you start on Thursday, you won't need to come in on the weekend.
:::


## Day 1: Seeding Cells

:::{.callout-note}
Prepare a suspension of your cells and count them
:::

```{r}
#| echo: false
#| column: margin

tibble("Cell Line" = c("1A6", "5637", "BV", "HT1197", "HT1376", "J82", "RT112", "RT4", "RT4v6", "SCaBER", "SW780", "TCCSUP", "UC1", "UC3", "UC6", "UC10", "UC14", "UC15", "UC16", "UC17", "UC18"), 
       "Cells/Well" = c(1000, 1000, 1000, 2000, 2000, 2000, 1000, 1000, 1500, 2500, 1000, 2000, 1000, 500, 350, 1500, 1000, 3000, 4000, 1000, 2000)) |> 
  gt() |> 
  tab_options(table.font.size = px(12))
```

1. Dilute the cells to appropriate concentration for your cell line (see table). Each well will contain $100\mu L$. You'll likely want to make one more mL of suspension than strictly needed so you can use a reagent reservoir and multichannel pipette.

2. Pour cell suspension into a reagent reservoir^[Cat# 8094, ThermoFisher], then use a multichannel pipette to dispense $100\mu L$ of suspension per well.


## Day 2: Adding Drug

1. Prepare a serial dilution of your drug. Prepare a DMSO control, which should contain the same volume of DMSO as the highest DMSO, drug-containing condition (this should be your highest drug concentration condition). 

:::{.callout-tip}
If you only need 1/4 of a 96 well plate for a given drug, you can prepare a serial dilution in 1mL microfuge tubes. Otherwise, use 15mL conicals.
:::

2. Aspirate medium from one 'condition' (usually 1/4 a 96 well plate)

:::{.callout-tip}
Use a filterless $1000\mu L$ tip on the typical 2mL aspirating tip to help aspirate medium from the wells
:::

:::{.callout-warning}
Try not to suck up the cells on the bottom. Tilt the plate towards you and aspirate on the bottom corner to limit then number of cells you aspirate.
:::

3. Use a P200 to pipette $100\mu L$ of drug medium per well.

:::{.callout-tip}
You can get away with using two tips per condition, so long as you aren't using these drugged media for other conditions. Start with the lowest (non-DMSO) drug concentration, then move to higher concentrations with the same tip. Finally, using another tip, add the DMSO medium.
:::


## Day 5: Replenishing Drug

1. Repeat Day 2, preparing new drug to replace the old medium.

:::{.callout-note}
Depending on the $t_{1/2}$ of your drug, this may need to replace the drug more frequently. Doxycycline, for instance, needs to be replaced roughly every other day.
:::

## Day 7: Measuring Metabolic Activity

1. Add a $1mL$ MTT aliquot to $11mL$ medium in a 15mL conical.

2. Aspirate medium from all wells of the 96 well plate.

3. Add $100\mu L$ of MTT solution to each well. Incubate for $1\ hr$ at $37^\circ C$. 


The following steps can be done outside the hood:

4. Aspirate MTT solution out of the wells. Add $50\mu L$ of DMSO to each well. Shake plate for $30\ sec$ in the SpectraMAX, then read at $562nm$ and $660nm$.

## Analysis and Interpretation

1. Subtract the 660nm absorbance (Background) from the 562nm absorbance (Wavelength that the MTT metabolite absorbs at) for each well
2. Average all technical replicates
3. Fit a dose response curve in Prism (R app coming soon)
