---
title: "MTT Analysis"
execute: 
  echo: false
---


:::{.callout-note}
This protocol assumes you have finished the [MTT Protocol](#sec-mtt-assay).

This protocol also requires you have installed R, RStudio, as well as the mop and bladdr package. bladdr is a GitHub package. Use `devtools::install_github("McConkeyLab/bladdr")` to install bladdr and `devtools::install_github("KaiAragaki/mop")` to install mop. For more information, see 'Installing GitHub Packages'.
:::

## Data Acquisition
1. In the Spectramax software, click 'File > Import/Export > Export...' and export the results as a .txt file to a USB drive

## Reading Data into R
1. Open RStudio
2. Press Shift + Ctrl/Command + N. This will open a new script.
3. Run the following in your script, with the path to your file replacing the `"dummy_mtt_spectramax.txt"`

```{r}
#| echo: true
library(bladdr)
library(mop)

my_spectramax <- read_spectramax("dummy_mtt_spectramax.txt")
```

You should now have a `spectramax` object that looks like this:
```{r}
my_spectramax
```


Now we need to calculate the proper values for plotting. This means we need to annotate our plate with our experimental conditions. To do this, we use `mtt_calc`.

`mtt_calc` takes three arguments. One is the `my_spectramax` you just created. The second is the names of the conditions of each 'sector' of your plate. The third is the drug concentrations within that sector.

The order of sectors can be seen in @fig-sectors.
```{r}
#| column: margin
#| label: fig-sectors
#| fig-cap: "Order of sector labelling in the `condition_names` argument"
library(gplate)
library(ggsci)
gp(wells = 96) |>
  gp_sec("Sector", 4, 6, labels = c("1", "2", "3", "4")) |>
  gp_plot(Sector, size = 9) +
  gp_mini_theme() +
  theme_tufte() +
  scale_color_npg()
```

In my data, I know I only used two sectors - sector 1 and 2.

Additionally, you need to annotate the concentration of drug. It all needs numeric - not text - so it needs to be the same units. Additionally, it moves from left to right in each sector (see @fig-conc).

```{r}
#| column: margin
#| label: fig-conc
#| fig-cap: "Order of drug concentration labelling in the `drug_conc` argument"
library(ggplot2)
gp(wells = 96) |> 
  gp_sec("Sector", 4, 6, labels = c("1", "2", "3", "4")) |> 
  gp_sec("Drug", 4, 1, labels = c("1", "2", "3", "4", "5", "6")) |> 
  gp_plot(Drug, size = 9) +
  gp_mini_theme() +
  theme_tufte() +
  scale_color_brewer()
```

All this being said, here's how I'll annotate my plate:

```{r}
#| echo: true

my_mtt_calc <- mtt_calc(
  my_spectramax, 
  condition_names = c("Drug 1", "Drug 2", NA, NA),
  drug_conc = c(0, 1, 10, 100, 1000, 10000)
)

my_mtt_calc
```

And these can be plotted like so:

```{r}
#| echo: true
mtt_plot(my_mtt_calc)
```

