# 3D Culture

```{r}
#| include: false

library(tidyverse)
library(gt)
```

## Handling Matrigel {#handling-matrigel}

* Matrigel should be aliquoted to avoid freeze-thawing (which can take quite a while).  

:::{.callout-tip}
Before aliquoting, it is often useful to spin the Matrigel to remove any solidified bits of Matrigel that may be present. If not done this can possibly impede clear imaging. 
:::

* Store aliquots at $–80^\circ C$ 

* Thaw on ice or in the refrigerator. For larger quantities, this make take overnight. 

* Any tubes that Matrigel is going to be added to should be chilled on ice. 

* Medium used to dilute Matrigel should also be chilled on ice. 


## Spheroids

### Forming Spheroids

:::{.callout-note}
Grow cells to $<80\%$ in a $T75$ (plenty for a 96 well plate)

Warm trypsin, medium, PBS

Add a glass bottom plate to plate warmer, set at $37^\circ C$.

Thaw Matrigel on ice. You'll need at minimum $100\mu L$ per well. Since it is highly viscous, thaw much more than necessary - it can always be refrozen if [handled properly](#handling-matrigel)
:::

1. Aspirate medium
2. Wash twice with $10mL$ PBS
3. Add $1.5mL$ trypsin. Regularly check the flask and tap against the palm of your hand.
4. Once $~80\%$ of the cells have dissociated, add $8.5mL$ medium
5. Transfer to a 15mL centrifuge tube.
6. Spin down at $300 \times g$ for $5\ minutes$
7. Resuspend in $10mL$ PBS
8. Spin down at $300 \times g$ for $5\ minutes$
9. Resuspend in $10mL$ medium and [count cells]($count-cells)
10. Prepare $12mL$ of cells diluted to the optimal cells/well concentration

:::{.callout-note}
If you are unsure of what concentration you should used, several concentrations should be prepared (250, 500, 1000, and 3000 cells/well is a good start)

Generally 500 cells/well works well

Each well should receive $100\mu L$ of volume
:::

11. Pour medium into a reagent reservoir. Use a multichannel pipette to dispense $100\mu L$ of volume to each well of a 96 well, round bottom, low-attachment plate (Cat. CLS7007-24EA)

12. Spin plate at $300 \times g$ for $5\ minutes$
13. Place plate in incubator

:::{.callout-tip}
If you're having difficulty getting good spheroid formation, you can try to layering $100\mu L$ $2\%$ Matrigel (for a final concentration of 1% Matrigel) medium on top of the wells after they have been spun down (being careful not to disturb the cells clustered at the bottom of the well). Remember to use [proper Matrigel handling technique](#handling-matrigel) when mixing medium and Matrigel.
:::

### Embedding Spheroids

:::{.callout-note}
Thaw matrigel on ice.
:::


## Organoids

Questions:

- When should OCM have matrix and when should it not?

- When making OCM with matrix, Should the OCM be very cool when made so that the matrix doesn’t immediately gel?

- How is 2.5% BSA prepared? 
- I assume 2.5% is weight/volume? g/mL? 
- How much to make at a time?
- Is it sterile filtered?

- How is Collagenase in DMEM/F12 with ROCK inhibitor prepared? 
- How much to make at a time? 
- Stable for how long? 
- Sterile filtered?

- When determining if a 24-well plate or T25 flask is more appropriate for culturing, what does a ‘large amount’ of tissue look like?

- How is tumor mass recorded? 
- Does it need to be sterile? 
- What vessel is it put in to measure mass?

- How do you wash patient tissue with OCM? 
- What detemines if the OCM used is with or without matrix? 
- How small is a small piece that determines if we use 5mL or 1mL? 
- Does using a smaller volume actually matter, or is it ok to use just a larger volume?

- When determining if TrypLE Express digestion should be done, what determines if we skip it altogether or route the unfiltered chunks in the filter to the TrypLE digestion? 
- Why choose one over the other? 
- More theoretical question – will this result in a selection bias if we don’t do it? 
- I assume these are combined with one another at some point? 
-  Which point?

- How much PBS to resuspend in?

- Why no ROCKi in final suspension?

- What happens if you overshoot the incubation time with TrypLE?

- Why should new scientists use single wells of 24 well plates?

- What does a ’24 well plate amount’ look like vs a ‘T25’ amount look like?



- What is the purpose of incubating OCM w matrix for at least 10 min? To warm medium?

- How much OCM w matrix for t25?

- Is it necessary to do all the ‘5mL for smaller tissues’ in cases other than the last suspension?

- I’m assuming the last OCM resuspension is OCM with matrix? Shouldn’t the last suspension volume depend (rather than just a flat 5mL?)

- HCM protocol [here](https://bioscience.lonza.com/lonza_bs/CH/en/download/product/asset/36719) it can be stored at 4 deg in the dark for a month. Do we heed this?

- How is primocin prepared? 
- What is it diluted in? 
- How much to make at once?
- Aliquot size? (should be in protocol, just need to remember to add when I write it up) 
- Sterile filtered?
- ROCK inhibitor preparation? 
- What is it diluted in? 
- How much to make? 
- Aliquot size? 
- Sterile filtered?



TODO:

- Anti-adherence solution Cat. #
- New figure for figure 1
- Syringe/bottle filter cat. #
- Include figure 2
- If we can get away with it, less 'while this is happening' business. People will figure it out on their own.

:::{.callout-note}
Coat all surfaces that will come in contact with tissue in 2.5% sterile BSA (tips, tubes, reservoirs, etc.) to avoid loss of sample due to sticking. To coat:

- Pipette tips: Pipette BSA up and down with the tip.

- Tube: Add BSA to the tube, then invert or vortex to coat the entire inside of the tube

- Reservoir: transfer some BSA to a reservoir, then use a multichannel pipette to coat the sides of the reservoir. 

After each instance, return BSA to stock vessel. BSA can be reused indefinitely if kept sterile.
:::

:::{.callout-note}
Be gentle with tissue to ensure cell clusters - rather than single cells - are recovered. 
:::

### Reagents

#### Matrix

Matrix can refer to either Corning Matrigel or Cultrex Basement Membrane Extract. What is used depends entirely on availability.

#### Hepatocyte Culture Medium (HCM) {#hcm}

```{r}
tibble(Reagent = c("HCM Hepatocyte Culture Medium BulletKit"), 
       Volume = c("500 mL"), 
       `Cat. #` = c("CC-3198")) |> gt(rowname_col = "Reagent")
```

1. Combine all SingleQuot Supplements with the Hepatocyte Basal Medium. Rinse each SingleQuot with medium to extract all of the supplement.
3. Shake to mix. Store at $4^\circ C$.

#### Organoid Culture Medium (OCM) {#ocm}

Make on an as-needed basis, within 24hrs of use. The protocol below is for a 10mL aliquot. Scale as needed.

```{r}
#| echo: false

tibble(Reagent = c("HCM",
                   "Charcoal Stripped FBS",
                   "10mM Y-27632 (ROCK inhibitor/ROCKi)", 
                   "100X Glutamax",
                   "50mg/mL Primocin", 
                   "Penicillin-Streptomycin 10,000 U/mL", 
                   "Matrix"), 
       Volume = c("10mL", 
                  "500uL", 
                  "10uL", 
                  "100uL", 
                  "20uL", 
                  "100uL", 
                  "400uL"), 
       `[Final]` = c("",
                     "5%",
                     "10uM",
                     "1X",
                     "100ug/mL",
                     "100U/mL",
                     "4%"),
       `Cat. #` = c("CC-3198", 
                    "F6765-100ML", 
                    "72304", 
                    "35050061", 
                    "NC9141851", 
                    "15140122", 
                    "354230/3533-010-02")) |> gt(rowname_col = "Reagent")
```

1. Thaw ROCKi and Primocin on ice
2. Combine all reagents in an appropriatly sized vessel. 
3. Sterile filter with either a handheld syringe and filter for volumes <span>$<50mL$</span> or a bottle filter for volumes <span>$>50mL$</span>

#### OCM with Matrix {#ocm-w-matrix}

Make on an as-needed basis, within 24hrs of use. The protocol below is for a 10mL aliquot. Scale as needed.

```{r}
#| echo: false
tibble(Reagent = c("OCM", "Matrix"), 
       Volume = c("10mL", "400uL"), 
       `[Final]` = c("","4%"),
       `Cat. #` = c("","354230/3533-010-02")) |> 
  gt(rowname_col = "Reagent")
```

1. Combine items in conical
2. Vortex
3. Incubate at $37^\circ C$ for at least $10\ minutes$.

### Tissue Dissociation

:::{.callout-note}
Prepare OCM
:::

1. Record tumor mass
2. Wash patient tissue pieces with $5mL$ [ocm](ocm)
3. Transfer tissue pieces into a well of a 6 well plate
4. Using sterile scissors and forceps, mince the tissue into a paste. Mincing should take several minutes.

:::{.callout-tip}
If you prefer, you can use sterile scalpels to mince the tissue
:::

5. Add $10mL$ of warm DMEM/F12 with collagenase and ROCKi to the paste

:::{.callout-tip}
If there is a small amount of tissue, use $5mL$ instead.
:::

6. Transfer the tissue suspension into a <span>$10mL$</span> conical tube
7. Incubate at $37^\circ C$ in the cell culture incubator for around $1\ hour$, gently inverting every $15\ minutes$.

:::{.callout-note}
Incubation time varies depending on tissue stroma content, with higher stromal content taking a longer time. Continue digesting until cell clusters are on average less than ~100 cells. Monitor progress via microscope.
:::

8. Centrifuge for $5\ minutes$ at $400\times g$
9. Aspirate supernatant.
10. Re-suspend in PBS.
11. Assess average cluster size under microscope. If the average cluster size is around 2-10 cells, skip 
11. Centrifuge for $5\ minutes$ at $400\times g$
12. Aspirate supernatant.
13. Add $5mL$ of TrypLE Express + ROCKi
14. Incubate at $RT$ for $3\ minutes$, gently inverting the tube.

:::{.callout-warning}
Don't overshoot incubation time.
:::

15. Stop trypsinization with $10mL$ Advanced DMEM/F12
16. Assess the amount of tissue clusters <100 cells under microscope. Determine how many wells of a 24 well plate should be used, or if using a T25 is necessary.
17. Centrifuge for $5\ minutes$ at $400\times g$. 
- While spinning, add $1mL$ of anti-adherence solution to each well deemed necessary of the 24 well plate, or $3mL$ of anti-adherence solution to a T25 flask. Incubate at room temperature for $10\ minutes$ 
- Additionally, make a sufficient volume of [OCM with matrix](ocm-w-matrix) to plate 1mL of cell suspension for each well, or 5mL for a T25 (determined previously).
18. Pass suspension through a <span>$100\mu m$<\span> cell strainer. A fair amount of leftover tissue is expected

:::{.callout-tip}
If you have extra Advanced DMEM/F12, pour up to $5mL$ over the strainer to get more suspension through.
:::

19. Spin strained suspension for $5\ minutes$ at $400\times g$. During this time, aspirate anti-adherence solution and replace with an equal volume of HCM.
20. Aspirate supernatant.
21. Re-suspend in the previously determined volume of OCM with matrix.
22. Aspirate HCM from receiving vessel
23. Add cell suspension to receiving vessel
24. Place in incubator.


### Organoid Care

Check organoids every day for morphologic changes, organoid formation, and adherence to the bottom of the plate. Change medium with fresh medium twice per week.
