---
fig-cap-location: margin
reference-location: margin
---

::: callout-note
This information can now be found at http://wiki.mcconkeylab.com/en/protocols/3d-culture

Future updates will occur there.
:::

# 3D Culture

Many protocols here are adapted from [Padmanaban et al. 2020](https://doi.org/10.1038/s41596-020-0335-3), and that can also serve as a reference for additional protocols not (yet) listed here.

```{r}
#| include: false

library(tidyverse)
library(gt)
```

## Handling Matrigel {#handling-matrigel}

* Matrigel should be aliquoted to avoid freeze-thawing (which can take quite a while)^[Before aliquoting, it is often useful to spin the Matrigel to remove any solidified bits of Matrigel that may be present. If not done this can impede imaging.].

* Store aliquots at $–80^\circ C$ 

* Thaw on ice or in the refrigerator. For larger quantities, this may take overnight. 

* Any tubes that Matrigel is going to be added to should be chilled on ice. 

* Medium used to dilute Matrigel should also be chilled on ice. 

 
## Spheroids

### Forming Spheroids

[This protocol was adapted from the AggreWell 400 (Stemcell, Cat# 34415) product guide (Stemcell, Document 10000001001 Ver. 3)]{.aside}

:::{.callout-note}
Grow cells to $~80\%$ in a $T75$ or $T175$

Warm trypsin, medium (complete), [base medium](#reagent-defition-glossary), and PBS
:::

#### Preparing the Plate

1. Add Anti-Adherence Rinsing Solution^[Stemcell, Cat# 07010] to each well of an AggreWell plate ($.5mL$ for 24 well, $2mL$ for 6 well)

:::{.callout-note}
There are four kinds of AggreWell plates - two well numbers (6 well and 24 well) each with two different microwell sizes within those wells (little divots that the cells form small clusters in) ($400\mu m$ and $800\mu m$ width). At the time of writing, it's not clear which is best for our purposes.
:::

2. Centrifuge for $5\ minutes$ at $1300 \times g$
  * Balance using a normal plate with water
  * If there are bubbles in the microwells after spinning, spin again.
3. Aspirate the anti-adherence rinsing solution
4. Add $2mL$ (24 well) or $5mL$ (6 well) basal medium to each well
5. Aspirate basal medium from each well
6. Add $1mL$ (24 well) or $2.5mL$ (6 well) complete medium to each well


#### Preparing the Cells

1. Aspirate medium from flask
2. Wash twice with $10mL$ PBS
3. Add $1.5mL$ trypsin. Regularly check the flask and tap against the palm of your hand.
4. Once $~80\%$ of the cells have dissociated, add $8.5mL$ medium
5. Transfer to a 15mL centrifuge tube.
6. Spin down at $300 \times g$ for $5\ minutes$
7. Resuspend in $10mL$ PBS
8. Spin down at $300 \times g$ for $5\ minutes$
9. Resuspend in $10mL$ medium and [count cells]($count-cells)
10. Prepare $10mL$ of cells diluted to the optimal cells/well concentration.
  * Calculate using the following formula, keeping in mind the end volume is $2\frac{mL}{well}$ for 24 well plates and $5\frac{mL}{well}$ for 6 well plates:
  
  $$target\frac{cells}{sphere} \times S\frac{sphere}{well} = required\frac{cells}{well}$$
  * $S$ stands for the expected number of spheroids per well (see margin table)
```{r}
#| column: margin
#| echo: false
library(tibble)
library(gt)
tibble(`# of Wells` = c(6, 6, 24, 24), 
       `Size of Microwells` = c("400um", "800um", "400um", "800um"), 
       `Expected Number of Spheroids` = c(5900, 1500, 1200, 300)) |> 
  gt()
```
  * If I wanted $500\frac{cells}{sphere}$ and I had 24 well AggreWell 400 plates, I would need:
  
  $$500\frac{cells}{sphere} \times 1200\frac{sphere}{well} = 600,000\frac{cells}{well}$$
  
:::{.callout-note}
The optimal number of cells per sphere depends somewhat on cell line and downstream application. 500 cells per sphere is a good place to start.
:::

10. Prepare a balance plate for centrifugation
11. Add the proper volume of cell suspension per well
12. Pipette to mix the wells, avoiding introducing bubbles
13. Spin down immediately, for $3\ minutes$ at $100 \times g$

:::{.callout-caution}
Set the centrifuge brake to 'off' to prevent splashing of sample out of plate
:::

14. Place plate in incubator $1-2\ days$

:::{.callout-tip}
  If you're having difficulty getting good spheroid formation, you can try to layering medium with Matrigel (for a final concentration of 1% Matrigel after added to the well) on top of the wells after they have been spun down (being careful not to disturb the cells clustered at the bottom of the well). Remember to use [proper Matrigel handling technique](#handling-matrigel) when mixing medium and Matrigel.
:::

### Feeding Spheroids

1. Slowly remove ~50% volume from the well and discard
2. Slowly add the same volume of medium, dispensing down the side of the well

### Harvesting Spheroids

:::{.callout-note}
If you plan to embed your spheroids, you should make sure you are prepared to do so soon after (eg thawing Matrigel)
:::

1. Pipette up and down once to suspend the spheres, then pull the suspension up and transfer to a 15 mL conical.
2. Spin down spheroids ($400 \times g$ for $5\ minutes$), or simply wait until a sufficient number of spheres have settled to the bottom of the tube.
3. Aspirate supernatant

:::{.callout-caution}
Do not repeatedly pipette up and down - you may break the spheres!
:::

### Embedding Spheroids

#### Matrigel Embedding

:::{.callout-note}
Thaw Matrigel on ice. This may take several hours. You will need $100\mu L$ of Matrigel per well, plus around 10% extra for inevitable loss.

Warm a glass-bottom plate on the heatblock in the tissue-culture room to $37^\circ C$
:::

1. Add $110\mu L$ of Matrigel per well to your spheroid pellet and resuspend^[Note that too many spheroids per well can make imaging difficult, and can affect paracrine signalling. Aim for ~125 spheroids per well in 24 well plates.].

2. Bring the heatblock with the glass-bottom plate into the hood and plug it in

3. Slowly pipette $100\mu L$ of spheroid suspension into each well, trying to aim for the center of the well and not have the liquid come in contact with the edges.

:::{.callout-note}
Hold the pipette tip slightly above the bottom of the well. The goal is to form a 'bulge' or 'dome' of Matrigel, not a layer.
:::

4. Place in incubator for $30\ minutes$.

5. Gently add $2mL$ medium to each well, ejecting slowly against the well wall.

#### Collagen Embedding

## Organoids

Questions:

- When should OCM have matrix and when should it not?

- When making OCM with matrix, Should the OCM be very cool when made so that the matrix doesn’t immediately gel?

- How is 2.5% BSA prepared? 
- I assume 2.5% is weight/volume? g/mL? 
- How much to make at a time?
- Is it sterile filtered?

- How is Collagenase in DMEM/F12 with ROCK inhibitor prepared? 
- How much to make at a time? 
- Stable for how long? 
- Sterile filtered?

- When determining if a 24-well plate or T25 flask is more appropriate for culturing, what does a ‘large amount’ of tissue look like?

- How is tumor mass recorded? 
- Does it need to be sterile? 
- What vessel is it put in to measure mass?

- How do you wash patient tissue with OCM? 
- What detemines if the OCM used is with or without matrix? 
- How small is a small piece that determines if we use 5mL or 1mL? 
- Does using a smaller volume actually matter, or is it ok to use just a larger volume?

- When determining if TrypLE Express digestion should be done, what determines if we skip it altogether or route the unfiltered chunks in the filter to the TrypLE digestion? 
- Why choose one over the other? 
- More theoretical question – will this result in a selection bias if we don’t do it? 
- I assume these are combined with one another at some point? 
-  Which point?

- How much PBS to resuspend in?

- Why no ROCKi in final suspension?

- What happens if you overshoot the incubation time with TrypLE?

- Why should new scientists use single wells of 24 well plates?

- What does a ’24 well plate amount’ look like vs a ‘T25’ amount look like?



- What is the purpose of incubating OCM w matrix for at least 10 min? To warm medium?

- How much OCM w matrix for t25?

- Is it necessary to do all the ‘5mL for smaller tissues’ in cases other than the last suspension?

- I’m assuming the last OCM resuspension is OCM with matrix? Shouldn’t the last suspension volume depend (rather than just a flat 5mL?)

- HCM protocol [here](https://bioscience.lonza.com/lonza_bs/CH/en/download/product/asset/36719) it can be stored at 4 deg in the dark for a month. Do we heed this?

- How is primocin prepared? 
- What is it diluted in? 
- How much to make at once?
- Aliquot size? (should be in protocol, just need to remember to add when I write it up) 
- Sterile filtered?
- ROCK inhibitor preparation? 
- What is it diluted in? 
- How much to make? 
- Aliquot size? 
- Sterile filtered?



TODO:

- Anti-adherence solution Cat. #
- New figure for figure 1
- Syringe/bottle filter cat. #
- Include figure 2
- If we can get away with it, less 'while this is happening' business. People will figure it out on their own.

:::{.callout-note}
Coat all surfaces that will come in contact with tissue in 2.5% sterile BSA (tips, tubes, reservoirs, etc.) to avoid loss of sample due to sticking. To coat:

- Pipette tips: Pipette BSA up and down with the tip.

- Tube: Add BSA to the tube, then invert or vortex to coat the entire inside of the tube

- Reservoir: transfer some BSA to a reservoir, then use a multichannel pipette to coat the sides of the reservoir. 

After each instance, return BSA to stock vessel. BSA can be reused indefinitely if kept sterile.
:::

:::{.callout-note}
Be gentle with tissue to ensure cell clusters - rather than single cells - are recovered. 
:::

### Reagents

#### Matrix

Matrix can refer to either Corning Matrigel or Cultrex Basement Membrane Extract. What is used depends entirely on availability.

#### Hepatocyte Culture Medium (HCM) {#hcm}

```{r}
#| echo: false

tibble(Reagent = c("HCM Hepatocyte Culture Medium BulletKit"), 
       Volume = c("500 mL"), 
       `Cat. #` = c("CC-3198")) |> gt(rowname_col = "Reagent")
```

1. Combine all SingleQuot Supplements with the Hepatocyte Basal Medium. Rinse each SingleQuot with medium to extract all of the supplement.
3. Shake to mix. Store at $4^\circ C$.

#### Organoid Culture Medium (OCM) {#ocm}

Make on an as-needed basis, within 24hrs of use. The protocol below is for a 10mL aliquot. Scale as needed.

```{r}
#| echo: false

tibble(Reagent = c("HCM",
                   "Charcoal Stripped FBS",
                   "10mM Y-27632 (ROCK inhibitor/ROCKi)", 
                   "100X Glutamax",
                   "50mg/mL Primocin", 
                   "Penicillin-Streptomycin 10,000 U/mL", 
                   "Matrix"), 
       Volume = c("10mL", 
                  "500uL", 
                  "10uL", 
                  "100uL", 
                  "20uL", 
                  "100uL", 
                  "400uL"), 
       `[Final]` = c("",
                     "5%",
                     "10uM",
                     "1X",
                     "100ug/mL",
                     "100U/mL",
                     "4%"),
       `Cat. #` = c("CC-3198", 
                    "F6765-100ML", 
                    "72304", 
                    "35050061", 
                    "NC9141851", 
                    "15140122", 
                    "354230/3533-010-02")) |> gt(rowname_col = "Reagent")
```

1. Thaw ROCKi and Primocin on ice
2. Combine all reagents in an appropriatly sized vessel. 
3. Sterile filter with either a handheld syringe and filter for volumes <span>$<50mL$</span> or a bottle filter for volumes <span>$>50mL$</span>

#### OCM with Matrix {#ocm-w-matrix}

Make on an as-needed basis, within 24hrs of use. The protocol below is for a 10mL aliquot. Scale as needed.

```{r}
#| echo: false
tibble(Reagent = c("OCM", "Matrix"), 
       Volume = c("10mL", "400uL"), 
       `[Final]` = c("","4%"),
       `Cat. #` = c("","354230/3533-010-02")) |> 
  gt(rowname_col = "Reagent")
```

1. Combine items in conical
2. Vortex
3. Incubate at $37^\circ C$ for at least $10\ minutes$.

### Tissue Dissociation

:::{.callout-note}
Prepare OCM
:::

1. Record tumor mass
2. Wash patient tissue pieces with $5mL$ [ocm](ocm)
3. Transfer tissue pieces into a well of a 6 well plate
4. Using sterile scissors and forceps, mince the tissue into a paste. Mincing should take several minutes.

:::{.callout-tip}
If you prefer, you can use sterile scalpels to mince the tissue
:::

5. Add $10mL$ of warm DMEM/F12 with collagenase and ROCKi to the paste

:::{.callout-tip}
If there is a small amount of tissue, use $5mL$ instead.
:::

6. Transfer the tissue suspension into a <span>$10mL$</span> conical tube
7. Incubate at $37^\circ C$ in the cell culture incubator for around $1\ hour$, gently inverting every $15\ minutes$.

:::{.callout-note}
Incubation time varies depending on tissue stroma content, with higher stromal content taking a longer time. Continue digesting until cell clusters are on average less than ~100 cells. Monitor progress via microscope.
:::

8. Centrifuge for $5\ minutes$ at $400\times g$
9. Aspirate supernatant.
10. Re-suspend in PBS.
11. Assess average cluster size under microscope. If the average cluster size is around 2-10 cells, skip 
11. Centrifuge for $5\ minutes$ at $400\times g$
12. Aspirate supernatant.
13. Add $5mL$ of TrypLE Express + ROCKi
14. Incubate at $RT$ for $3\ minutes$, gently inverting the tube.

:::{.callout-warning}
Don't overshoot incubation time.
:::

15. Stop trypsinization with $10mL$ Advanced DMEM/F12
16. Assess the amount of tissue clusters <100 cells under microscope. Determine how many wells of a 24 well plate should be used, or if using a T25 is necessary.
17. Centrifuge for $5\ minutes$ at $400\times g$. 
- While spinning, add $1mL$ of anti-adherence solution to each well deemed necessary of the 24 well plate, or $3mL$ of anti-adherence solution to a T25 flask. Incubate at room temperature for $10\ minutes$ 
- Additionally, make a sufficient volume of [OCM with matrix](ocm-w-matrix) to plate 1mL of cell suspension for each well, or 5mL for a T25 (determined previously).
18. Pass suspension through a <span>$100\mu m$</span> cell strainer. A fair amount of leftover tissue is expected

:::{.callout-tip}
If you have extra Advanced DMEM/F12, pour up to $5mL$ over the strainer to get more suspension through.
:::

19. Spin strained suspension for $5\ minutes$ at $400\times g$. During this time, aspirate anti-adherence solution and replace with an equal volume of HCM.
20. Aspirate supernatant.
21. Re-suspend in the previously determined volume of OCM with matrix.
22. Aspirate HCM from receiving vessel
23. Add cell suspension to receiving vessel
24. Place in incubator.


### Organoid Care

Check organoids every day for morphologic changes, organoid formation, and adherence to the bottom of the plate. Change medium with fresh medium twice per week.
